---
title: Integrative cross-omics and cross-context analysis elucidates molecular links underlying genetic effects on complex traits
author: "Yihao Lu, Meritxell Oliva, Brandon L. Pierce, Jin Liu, and Lin S. Chen"
output:
  pdf_document: default
  latex_engine: xelatex
  word_document: default
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
This vignette provides an introduction to the `X.ING` package. The R package `X.ING` implements the X-ING method for cross-omics and cross-context analysis.

To install the development version of the `X.ING` package, please load the `devtools` package first. Note that `X.ING` requires the `CCA`, `Rcpp`, and `RcppArmadillo` packages. 

To install this package, run the following command in R
```
library(devtools)
install_github("ylustat/X.ING")
```

Load the package using the following command:
```{r}
library("X.ING")
```
This vignette depends on R packages `MendelianRandomization`, `ggplot2` and `tidyverse`, you can load these packages using the following command:
```{r}
suppressMessages(library("tidyverse"))
library("ggplot2")
library("MendelianRandomization")
```


### Fit X.ING for correlated SNPs using simulated data
In this section, we fit X.ING using simulated data (provided in the package) as an example to illustrate the basic usage of X.ING. `gamma_hat` and `se_g` are the estimated SNP-to-expression effect and its standard error. `LD` is the estimated linkage disequilibrium matrix among IVs. `latent status` is the true underlying effect status, where `1` indicates the presence of non-zero effect and `0` indicates the absence of effect.

