---
title: Integrative cross-omics and cross-context analysis elucidates molecular links underlying genetic effects on complex traits
author: "Yihao Lu, Meritxell Oliva, Brandon L. Pierce, Jin Liu, and Lin S. Chen"
output:
  pdf_document: default
  latex_engine: xelatex
  word_document: default
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
This vignette provides an introduction to the `X.ING` package. The R package `X.ING` implements the X-ING method for cross-omics and cross-context analysis.

To install the development version of the `X.ING` package, please load the `devtools` package first. Note that `X.ING` requires the `CCA`, `RGCCA`, `Rcpp`, and `RcppArmadillo` packages. 

To install this package, run the following command in R
```
library(devtools)
install_github("ylustat/X.ING")
```

Load the package using the following command:
```{r}
library("X.ING")
```
This vignette depends on R packages `MendelianRandomization`, `ggplot2` and `tidyverse`, you can load these packages using the following command:
```{r}
suppressMessages(library("tidyverse"))
suppressMessages(library("pROC"))
suppressMessages(library("MASS"))
library("ggplot2")
library("MendelianRandomization")
library("RGCCA")
```


### Fit XING using simulated data
The main function of the `X.ING` package is `XING()`.  In this section, we fit `XING` function using simulated data (provided in the package) as an example to illustrate the basic usage of `XING`. The example data is a list with two elements, where `z` is a list of z-scores and `gamma` is a list of true underlying association status. `1` in `gamma` indicates the presence of non-zero effect and `0` indicates the absence of effect. We have three sets of data available in the example data (L=3). The number of contexts are 10, 8, and 6, respectively.
```{r}
data("example_data")
length(example$z)
sapply(example$z, dim)
```
The proportion of true associations are ~0.04.
```{r}
sapply(example$gamma, mean)
```
The X-ING takes the z-score list as input. It automatically use CCA (when L=2) for GCCA (when L>2) for the multi-view learning. Afterwards, it applies PCA to extract shared patterns across contexts for each omics. A simplest application where two sets of z-scores are analyzed:
```{r}
z <- example$z
res <- XING(z_list = z[1:2])
```
The output includes posterior probabilities (`alpha`) and posterior means (`mu`):
```{r}
names(res)
lengths(res)
```
AUC can be calculated based on the posterior probability and true association status
```{r}
gamma <- example$gamma

mapply(function(g,pp) auc(c(g),c(pp), quiet = T), gamma[1:2], res$alpha)
```
It switches to GCCA (implemented in RGCCA package) if L>2 (note that the computation speed may substantially slowed):
```{r}
z <- example$z
res <- XING(z_list = z)
```
Similar to the previous example, the output also includes posterior probabilities (`alpha`) and posterior means (`mu`):
```{r}
names(res)
lengths(res)
```
And the corresponding AUCs:
```{r}
mapply(function(g,pp) auc(c(g),c(pp), quiet = T), gamma, res$alpha)
```
Users may specify the sample overlap matrix $\widehat{\boldsymbol{R}}$ through thr `Lambda_list` parameter. The number of canonical coefficients to keep, and the number of PCs can be specified by `CC` and `PC_list`. For example,
```{r}
res <- XING(z_list = z, CC = 4, PC_list = c(4,3,2), Lambda_list = lapply(sapply(z,ncol),diag))
mapply(function(g,pp) auc(c(g),c(pp), quiet = T), gamma, res$alpha)
```
To reduce computation time, user may specify the iteration time via `iterT`. Also, the `tolerance` parameter controls the criterion for convergence of the GCCA algorithm. A smaller `tolerance` may lead to substantially longer computation time.







